\newcommand{\wineIntro}{Tonight, we are grateful for the unprecedented freedom in our present day lives---far more freedom than our ancestors could ever have dreamed was possible.}

\newcommand{\wineSummaryThisYear}{This year, we will hear some strategies\textemdash{}adapted from David Cain's writings on Raptitude\textemdash{}to feel deep gratitude for the freedoms we enjoy, that we may better appreciate the importance of bringing freedom to the rest of the world.}

% Gratitude: how to be thankful for all the freedom you have

% first cup--notice what you have, it's not all about the huge things...tiny things are important too: https://www.raptitude.com/2018/02/gratitude-noticing/

% second cup: life is outrageously awesome: https://www.raptitude.com/2011/01/a-day-in-the-future/  https://www.raptitude.com/2013/10/use-your-privilege/

% 3rd cup (after grace): "It might not be so"--think about things being gone https://www.raptitude.com/2020/02/how-to-create-gratitude/

% 4th cup: people used to have it BAD---we have it really good https://www.raptitude.com/2016/02/a-brief-visit-to-the-end-of-the-world/
% https://www.raptitude.com/2013/10/use-your-privilege/

% https://www.raptitude.com/2011/01/a-day-in-the-future/  maybe mix in https://www.raptitude.com/2011/09/how-to-walk-across-a-parking-lot/  or maybe mix in https://www.raptitude.com/2013/10/use-your-privilege/




% the "Peacetime Dream": https://www.raptitude.com/2016/02/a-brief-visit-to-the-end-of-the-world/

% ...probably not: https://www.raptitude.com/2015/06/you-are-free-like-it-or-not/


% generic war-based freedom...not great https://bmcnews.org/story/an-appreciation-for-freedom
% more war things https://www.dialogueworks.com/blog/what-freedom-do-you-value-most-seven-freedoms-that-you-may-take-for-granted
% https://www.dialogueworks.com/blog/8-ways-to-appreciate-our-country-

\newcommand{\wineFirstCup}{
  % Typically takes up about 1.5 pages. runs into the 2nd half of Page 3...leaving some space, but bumping the hebrew onto the second page
  \newReading With this first cup of wine, we consider how to feel more grateful for the freedom we enjoy 
  %not by just hearing about the lack of freedom others have, but 
  by noticing the experiences we have that happen because we are free.
  
  Many gratitude exercises ask you to think about what you have to be grateful for. In other words, you brainstorm reasons you ought to feel grateful, whether or not you do.

  These exercises are worthwhile on some level, but most of the time they don't create much of a real-time, visceral sense of gratitude. They just remind you of certain encouraging rote facts: on paper, your situation is pretty good; many parts of your life would be enviable to others; our ancestors certainly had things far worse as slaves in the Land of Egypt.

  \newReading As you might have noticed, simply making the case to ourselves that we have \textit{reasons} to feel grateful and celebrate our freedom doesn't necessarily make us feel that way.
  
  Gratitude, when we do genuinely feel it, arises from experiences we are \textit{currently having}, not from evaluating our lives in our heads. When you feel lonely, simply remembering that you have friends is a dull, nominal comfort compared to how wonderful it feels when one of those friends calls you out of the blue. Reflecting on the good fortune of having a fixed address is nice, but stepping inside your front door after a cold and rainy walk home is sublime. 
  
  The experience, not the idea, is what matters. So if you want to feel grateful for your freedom, move beyond the thinking exercises. Look for your good fortune not in some abstract assessment of your life situation, but in your experience right in this moment. What can you see, feel, hear, or sense, right here in the present, that's helpful, pleasant, or beautiful? Any interesting sensory experience or pleasant feeling will do: the cool breeze from a fan, the pet on your lap, the taste of this wine.
  
  \newReading That's the other important part: we don't need to reserve our gratitude for the big, fundamental conditions, % TODO (Eli 20220403): come up with better adjective for all-encompassing
  such as health, economic class, or loved ones. In every moment, regardless of your problems, you are enjoying freedom through all sorts of small, often haphazard pleasures: the smell of a meal wafting in from the kitchen, the softness of this chair, the alertness of your mind right now.
  
  Of course, all the abstract, big-picture life conditions made possible by your freedom have their own corresponding present-moment pleasures, and they are what matter. Consider the world of difference between trying to appreciate the notion that you aren't homeless, and appreciating the real-time experience of getting into bed in your own bedroom. That's where your good fortune truly resides---in your experiences, not your thoughts.
  
  With this first cup of wine, we remember to enjoy our present experiences to help us feel grateful for all the freedom that has been given to us.
}

\newcommand{\wineSecondCup}{
  % About 1.5-2 pages
  \newReading With this second cup of wine, we learn a technique to more easily feel gratitude during everyday experiences; during a seemingly ordinary activity, describe it in a way that would be understandable to our ancestors in Egypt. 
  
  %try to describe it to our ancestors in Egypt. and from the perspective of trying to describe it to our ancestors in Egypt.
  
  %look at our lives from the perspective of trying to describe them to our ancestors in Egypt to help us better appreciate the incredible freedoms we enjoy.
  
  I wake up in a soft bed. I'm warm and safe, like every morning. %Outside it is below freezing, but from inside my home winter seems far away.
  As I rise and stretch, I notice I'm sore. Not from tending the fields though. I have no fields. %Some unseen person does all the field-tending for me. 
  Sometimes I forget that there's any field-tending going on at all. I buy all my food---I wouldn't know how to grow it or hunt it. I've never gone a day without food.
  
  My soreness is actually from my leisure time, not work. I spent yesterday sliding down a snow-covered slope with a board attached to my feet. After that I was pretty worn out, so I went to a friend's house, and watched a sporting event as it unfolded in Boston. I don't live in Boston, but my friend has a machine that lets us see what's happening there. I have one too. Almost everyone does.
  
  \newReading The sun won't rise for another hour, but I don't need to light a fire or candles. I have artificial ones, mounted on the ceiling. Hit a tiny switch and I can see everything, any time of day.
  
  I bathe while standing. The water comes out whatever temperature I like.
  
  I use a few machines in my kitchen to get my breakfast ready. It takes about five minutes. %Toasted buckwheat groats with raisins, almonds, dates and sunflower seeds. 
  I don't know where it came from but I'd be surprised if it was from anywhere near here. As I eat, it occurs to me that my co-worker has a machine I might need to use at work today, so I want to make sure he brings it with him. He lives five miles from me, but that's no problem. I've got a device that lets him hear my voice from that distance. Wherever he is, I can talk to him.

  One minute later I've solved that problem, and forgotten about it.

  \newReading I put my dishes into another machine, which will clean them for me while I'm away at work.
  %I get dressed and leave. 
  My destination is ten miles in the distance and I've got twenty-five minutes, which is plenty of time, because I won't be walking.

  I get into my most expensive machine. It's actually quite miraculous, but I forget that all the time. It allows me to sit in a comfortable chair, sealed from the elements, while it propels me at incredible speeds.

  %Just like my home, I can make it any temperature I wish inside. I don't have to exert any real effort to make the thing go. I use my hands and my toes to control it.

  %I don't know quite how it works, but it's powered by a liquid I can buy pretty much anywhere. I can transport hundreds of pounds of whatever I want, and even listen to long-dead musicians singing and playing instruments while I do it.

  %So I hurtle myself to my workplace, which is well beyond the horizon, looking from my house. There, I do what a corporation asks me to, for most of my daylight hours. It's not that tough really.

  After work, I hurtle home in the same manner, without really thinking about it. %I make dinner for myself, and eat. 
  Then I turn on one of my favorite machines. It's about the size of a scrap of parchment.

  It has a glowing window inside it. A single page. But I only need one page because its contents change at my command. Sometimes there are words, sometimes pictures, sometimes both. Some pictures can move and talk.
  
  \newReading The stuff inside can be written by anyone in the world, even \textit{while it's in my hand}. There's more inside than I could ever read. It provides me with unbelievable advantages. Anything I don't know, I can find out in a few seconds.
  %I can get instructions on how to do pretty much anything that has ever been done. I can summon complete histories of almost any person or culture you could name, expert opinions on anything at all, unlimited advice, unlimited entertainment, unlimited information. 
  I can buy pretty much anything from where I'm sitting, and have it brought to my door.
  
  The sun went down hours ago, but with my artificial light I haven't noticed. When I'm able to summon the willpower, I close my favorite machine and go to bed.

  I'm warm and safe, like every night.
  
  We drink this second cup realizing that our ancestors would not have been able to even conceive of a life filled with as much freedom as we enjoy today. 
}

\newcommand{\wineThirdCup}{
  % should be 1.5-ish pages
  \newReading With the third cup of wine, we appreciate the tremendous benefits in our lives afforded by our freedom by imagining if they were to disappear.
  
  As a kid, whenever I stayed for supper at certain friends' houses, I wasn't sure what to do when they prayed.

  My family didn't say grace, but I knew a bit about the ritual from TV. I knew you were supposed to look down and say amen at the end, so I did.
  
  %I was familiar with the idea of \god{}---how he made the world and watched over it, and all that. But I found it unlikely he would intervene in the pedestrian matters of cooking and groceries. %Still, it made as much sense as Santa Claus and the impossible logistical feats attributed to him, so I went through the motions in the way kids do.
  
  ``Grace isn't really religious,'' a formerly religious co-worker said to me years later. %We had picked up junk food to eat on the drive to the next jobsite. It was probably the unhealthiest meal possible---energy drinks and 7-Eleven butter tarts---and I had joked that we should say grace.
  
  ``Yeah, you don't have to believe the food came from the Lord. All they're saying is, `Before we eat this food, let's remember that we could have no food.'{}''
  
  \newReading \textit{This might not be so}, in other words. This thing you have because you are free and not a slave in Egypt---this meal, this bed, these clothes, this friend---if it's possible to have it, it's possible not to have it. If you take a moment to imagine not having it, the good fortune of having it is no longer lost on you.
  
  Some nights, once I was comfortably in bed, I would imagine my bed disappearing around me and dropping me onto the bare floor. A moment later, the whole house would go, depositing me onto the cold, wet grass.
  
  I'd imagine this scenario in vivid detail. The dirt and dampness, the dead grass stuck to my arms. After a minute or so, the daydream would break, and the warmth and comfort of the covers---the happy reality of where I really was---would hit me like a truck.
  
  \newReading I'd do this on a smaller scale too, such as imagining my socks disappearing, just long enough to feel my clammy bare feet against the damp insoles of my boots. A moment later, I'd drop the fantasy, and discover anew just how much comfort a cheap pair of sport socks was adding to my life in that moment.
  
  I still do this. With socks, sweaters, walls. It's just one way to get your mind to collide with the essential truth behind any genuine experience of gratitude: this might not be so.  
  
  If you do the same exercise not with socks but with your loved ones, the effect is even more powerful. You'll need a moment in which you're with a loved one, and they're talking to someone else, doing a crossword, or otherwise not directly engaging with you.
  
  \newReading Whatever they're doing, quietly observe them doing it, paying attention to their unique mannerisms, voice, and presence.
  
  %Here's the part that will make you cry. 
  Now, try to see this moment as though it's actually in the past, and this person is gone now. You're remembering this one ordinary moment, in perfect detail, from that precious window of time when your lives still overlapped.
  
  It's important that it's an ordinary moment, because those are the kinds of moments that often seem neutral in value. Abundant and unremarkable. You might be tempted to spend it checking your phone.
  
  But if your mind can find that place, even for a just a second, where you see that \textit{this might not be so}, you'll break that sense that nothing special is happening.
  
  \newReading Then there will be a moment when you come back to reality. And the reality is that the precious, fleeting time when you got to be near this person\textemdash{}the time you'd do anything to revisit\textemdash{}is happening right now.
  
  The payoff isn't in the morbid fantasy about loss, it's the instant when you return from the vision and recognize what you really have, that you're living right in the middle of a fragile golden era. It's unique and beautiful and will one day be gone, and you are fantastically lucky to be here for it.
  
  With this third cup of wine, we feel gratitude for all of our freedoms and the joy they bring to our lives, by remembering that \textit{it might not be so}.
}

\newcommand{\wineFourthCup}{
  \newReading As our Seder draws near to the end, we take up our cups one last time as we think about how to feel grateful for all the freedom we enjoy even when faced with something unexpected.
  
  ``Radical Gratitude'' is about experimenting with being grateful for everything that happens to you. You're not telling yourself you \textit{should} feel grateful, only to invite or explore gratitude for what happens regardless of our initial feeling about it. The idea sounds ridiculous, and even hopeless, but in practice it's quite easy, and immediately rewarding. You just ask yourself, ``Can I be grateful for this too?'' In my short experience doing this, the brain has a way of coming up with good reasons why yes, you can.

  This practice reveals a lot about our short-sightedness. We have a rather ridiculous tendency to believe everything is either strictly good for us or bad for us, and that we can reliably determine which one it is, in the instant that thing happens.
  
  \newReading A few weeks ago, I went down to my storage locker and saw the door hanging open. My bike was gone. My initial feeling was the rush of violation and dirtiness that everyone feels when they see the mess left by a thief. They touched my stuff, and now some of it is at their place. But then I asked myself, ``Is there anything about this I could be grateful for?''
  
  The thought immediately put me into a totally different position, one where I didn't assume I should feel any particular way about it. The normal victim feelings gave way to something different. %a feeling of, ``Wow, I'm really glad I'm me.'' 
  I'm glad to know that the locker was so insecure before I put anything irreplaceable in there. I can afford a new bike. Also, I've never felt a desire to steal from people. Aren't I lucky that I don't know what it's like to enter a building illegally, and rifle through someone else's belongings? 
  
  %I wasn't thinking of it at the time, but I had recently listened to a short talk about cultivating gratitude at unusual moments. 
  

  % I live in the city and make use of street parking every day. Most of the time I can't find a spot on the stretch near my building, and I have to go past the building around to the long side of the block. Predictably I curse my bad luck, and often the people who had parked there inconsiderately, or at all.

  % Just after I'd listened to the talk, this happened to me again. I was on the cusp of re-enacting my normal sequence of overreaction\textemdash{}disappointment, maybe rage, then grumpy trudging\textemdash{}when I remembered the practice. Could I be grateful that I couldn't find a spot close to the building\textemdash{}that what's happening is in many ways a good thing?

  %The thought immediately put me into a totally different position, one where I didn't assume I should feel any particular way about it.

  % Mostly I just enjoyed the walk, noticed a few of my neighbor's yard decorations, and felt glad that carrying grocery bags two blocks isn't particularly difficult for me. I'm lucky to be able to walk almost any distance without chronic pain or fatigue. It struck me that my neighborhood is so close to downtown yet is really peaceful and safe. I can walk through it at 4am with nothing to worry about.
  
  %I arrived at my door feeling rather thrilled with my position in life, for exactly the way things were unfolding right now.
  
  \newReading Radical Gratitude is simply a way of challenging our initial feeling that a new development is wholly bad and that our moping and anger is justified, exploring instead what might also be good about it. Primarily, it does two things:

  It forces us out of the hypersensitive kind of autopilot we often operate under, which is based on a pretty grievous misconception: that events are isolated and are of two distinct types\textemdash{}good or bad\textemdash{}and that this goodness or badness is determined by how welcome it \textit{feels} when it happens.

  It also puts you into a helpful problem-solving state that often ends in gratitude for \textit{something} about what has just happened\textemdash{}the doors it opens, the things it teaches you, the future trouble it might spare you.

  \newReading Experimenting with this is also kind of fun. The more absurd it seems to be grateful, given the situation, the more interesting and fun it can be. Can I be grateful that my plans were canceled? Certainly. Can I be grateful I have a rash on my foot? Uh, let's find out.
  
  Again, being grateful for \textit{everything} is not an obligation; it's an option to experiment with. You aren't responsible for feeling any particular way after something happens.

  But that initial feeling\textemdash{}the horror of a locker door hanging open\textemdash{}doesn't have to be the final word on what's good and what's not. It's a question too complicated for our most reactive emotions to answer.

  With this fourth cup of wine, we consider how much easier it might be for us to find something to be grateful for in any event we're confronted with, compared to life for our ancestors in Egypt and those around the world still living in slavery today.
}

% \newcommand{\wineFourthCup}{
%   \newReading As our Seder draws near to the end, we take up our cups one last time as we think about how unique the amount of freedom we enjoy in our lives today is compared not just to our ancestors in Egypt, but every time before the Exodus and since.
  
%   People mostly want the same thing, and many of us already have it today, but we don't really notice it.

%   I have no way of confirming this, but I bet that if you could interview people across different centuries and cultures, asking them what they wanted most, you would notice a distinct theme in their answers.

%   Some people would want great riches or power. Others would say they want something very specific: to invent a particular thing, or for a particular person to love them, or to win a gold medal or give an Oscar speech.

%   But I suspect most of them would say they want something like this:
  
%   \noindent\makebox[\textwidth][c]{
%     \begin{minipage}[t]{0.9\textwidth}
%       \textit{I want to be able to do my work and spend time with my friends and family, free to live my own values in relative peace. I just want a fair chance to pursue love and happiness, and a stable, humble life.}
%     \end{minipage}
%   }  
  
%   You could call this the ``Peacetime Dream'', a life with the normal share of ups and downs---necessarily including heartbreak, health issues, setbacks and disappointments---but which isn't defined by war or persecution. Almost universally, people want basic stability and basic freedom, and to someone who doesn't have those things they are clearly the best things in the world.
  
%   But to someone who does have those things, their greatness is not so clear. It's easy to forget, or never notice at all, that many or most of us at this table already have this state of affairs, more or less.
  
%   It's also easy to forget that most of Jewish (and indeed all humankind's) history never had the Peacetime Dream. I wonder how many billions of individual human lives have been lived under slavery, tyrannical regimes, or during a war or a plague, or maybe all of those things.
  
%   Imagine arranging the lives of every person who has ever lived, in order of how much freedom they had in their time, to do the kind of work they wanted to do and live the kind of life they wanted to live. Where do you think you fit? It's extremely unlikely that you aren't near the very top. %Argue with that at your peril.
  
%   I don't know what your everyday worries are about, but I often worry about things like my work being criticized, the difficulty in making friends post-college, the ease of putting on weight at Thanksgiving, the advance of age, the murkiness of our tax laws, and the declining quality of consumer products. 
  
%   One defining characteristic of all of these concerns, aside from their ability to dominate my mind on a regular basis, is that they would evaporate the instant my basic freedom and stability were threatened. If the power went out and it became clear that it wasn't coming back on, suddenly few of our day-to-day worries would matter much. All we would care about is regaining the Peacetime Dream, that relatively stable state of affairs that allows a person to build a life.
  
%   But when we live in the Peacetime Dream, and always have, we don't even notice it. Imagine already having the one thing human beings most covet, and not knowing it. Our ability to take this foundation for granted is quite amazing. Even when I do worry about true horrors like slavery, starvation and violence, I'm worrying on behalf of other people. I believe (without realizing it) that it will always be someone else who experiences them.
  
%   In those rare moments when you're aware of the Peacetime Dream, to even go for a walk down the street is a joy. I remember being touched by a photo of a British Spitfire pilot crouching down to kiss the ground after a rough flight. Even if it's right under our feet, we simply can't see the value of what we have until we have some sense of being without it.
  
%   Imagine, for a moment, that the internet-using class of the Western world is, right now, experiencing the peak of human civilization---that there will never be more freedom, for the typical citizen, for making a great life than there currently is. The typical Westerner has never been more privileged, never had such a well-paved path to personal greatness and fulfillment. And imagine that during an era of incredible technology and possibility, most of these privileged people continued to overspend, underachieve, complain, and fill their time with minimal-return activities like watching reality television.
  
%   If this year were the peak of civilization, and in a few decades people looked back on the era we're in as The Golden Age, the younger generation would view many of us the way we view lottery winners who end up broke. They would find it hard to believe you were waiting for a better time.
  
  
%   With this fourth cup of wine, we are grateful that we are living the Peacetime Dream, and for our unique place in history with more freedom in our lives than ever previously existed.

% }