\newcommand{\wineIntro}{But we also reflect on what ``freedom'' means to
different people and what it means to each of us.}

\newcommand{\wineSummaryThisYear}{This year, we think about different views of
freedom.}

\newcommand{\wineFirstCup}{
  \newReading With this first cup of wine, we compare how some individuals
  define their own view of freedom.

  

  \subsubsection*{What is freedom?}
  What is freedom? How is freedom related to liberty? Liberty conveys the idea
  of the right or the ability of the individual to do what he wants. Freedom on
  the other hand is a word which contains inbuilt restrictions. 

  My freedom to swing my hands stops at my neighbor's nose. A free society
  contains individuals who are free to do what they do, whilst respecting the
  freedoms of others. Freedom is the liberty of the individual coupled with
  concern by the liberated individual for the liberties of others.

  Freedom cannot be defined, except through an analysis of the restrictions on
  human action.
  \textemdash{}Dr. Mark Cooray

  \subsubsection*{Freedom's just another word for nothin' left to lose}
  \newReading Sure, having nothing to lose is a type of freedom. But it's a
  temporary type; freedom is certainly not synonymous with having nothing to
  lose. And if you really, truly have nothing at all to lose, you're not even
  remotely free; you're closer to slavery than freedom, since you'd better get
  cracking to find yourself something to eat within the next 24 hours or so.

  Real freedom, it seems to me, comes when you don't have to do what others tell
  you. Granted, possessions always bring responsibilities, but carefully chosen
  possessions can provide more freedom than they take away.

  I've always thought a better statement is: ``Freedom means never having to say
  `I'm sorry' to someone who's really a total jerk.''
  
  \textemdash{}Posted by ``jackelope'' on the StraightDope message boards 

  \subsubsection*{Free to be yourself}
  \newReading The most important kind of freedom is to be what you really are.
  You trade in your reality for a role. You trade in your senses for an act. You
  give up your ability to feel and in exchange, put on a mask. I think people
  resist freedom because they're afraid of the unknown.

  The only solution is to confront yourself with the greatest fear imaginable.
  Expose yourself to yourself and to your deepest fear. After that, fear has no
  power, and fear of freedom shrinks and vanishes. You are free.

  \textemdash{}Jim Morrison

  With this first cup of wine, we begin to compose our own definition of
  freedom. 


}

\newcommand{\wineSecondCup}{
  \newReading With this second cup of wine, we contemplate different ways to
  view freedom in our lifestyles.

  \subsubsection*{The Monk and the Minister}
  Two close boyhood friends grow up and go their separate ways.  One becomes a
  humble monk, the other a rich and powerful minister to the king.

  Years later they meet.  As they catch up, the minister (in his fine robes)
  takes pity on the thin, shabby monk.  Seeking to help, he says: ``You know, if
  you could learn to cater to the king you wouldn't have to live on rice and
  beans.''

  To which the monk replies: ``If you could learn to live on rice and beans, you
  wouldn't have to cater to the king.''

  \newReading While most all of us fall somewhere between these two examples,
  they provide a spectrum on which to consider freedom. 

  Forming business deals or other contracts, and making family or personal
  promises are actions we take freely that can bring great happiness, but at the
  same time they limit our freedom to follow other pursuits by binding us to
  those obligations. 

  For this second cup of wine, we ask ourselves: Is my view of freedom closer to
  the monk, or to the minister? And are my choices and actions leading me
  towards or away from that goal?
}

\newcommand{\wineThirdCup}{
  \newReading With the third cup of wine, we think about how different countries
  view freedom and how history may have impacted their cultural beliefs.
  
  \subsubsection*{Freer\textemdash{}and Less Free}
  Untrammeled freedom lifts the American spirit. For centuries, an American's
  imagination has always stood as some new frontier. It was America's Manifest
  Destiny to spread west across the continent to claim what seemed like
  limitless resources. The California Gold Rush drew thousands hoping to strike
  it rich and many still flock to Silicon Valley today for that same
  opportunity. To an American, it is an egregious incursion on freedom to be
  prevented from taking a risk and hoping that they will be among the few who
  defy the laws of chance to rise from the common ranks to the realm of untold
  riches.
  
  However, this liberty unchecked by solidarity does not make a European heart
  beat faster. There, the state is cherished as protector rather than reviled as
  predator. It is seen as the balancer of economic opportunity, not the brake on
  it. While the American borders underwent massive expansion, Europe's remained
  largely fixed. The European pie, being of static size, creates a situation
  where an expansion of one piece only comes at the expense of another. This
  drives government attempts to divide the pie in a fraternal manner in order to
  make the world more just.

  \textemdash{}NY Times

  \subsubsection*{Stability is Freedom}
  \newReading China is often exemplified as the largest body of unfree people in
  the world, but in some respects that reflects our view of freedom rather than
  theirs. Within China, many polls find that societal stability ranks as a much
  greater concern than unfettered free speech and the right to protest. They
  often see those ``freedoms'' undermining stability in America and Europe.

  It is important to note that of the ten bloodiest conflicts in world history,
  two were world wars, but five of the other eight took place or originated in
  China. The scale of the slaughter within a single country, and the frequency
  with which the place has been bathed in blood, is hard for other nations to
  comprehend. The Taiping revolt in the mid-19th century led to the deaths of
  more than 20 million, and a decade later conflict between Han Chinese and
  Muslims killed another 10 million. In the 20th century 25 million died under
  Mao Zedong: some murdered, most as a result of a famine caused by brutality
  and incompetence.

  China's Communist Party leaders are no doubt keen to hold on to power for its
  own sake. But the country's grim history also helps explain why they are so
  determined not to give up ground to groups of pro-democracy demonstrators.

  Stability provides freedom from turmoil and death, which may be worth the
  sacrifice of other freedoms for long into the future in the minds of many
  Chinese. 
  
  \textemdash{}The Economist
  
  \newReading Are these different views of freedom a logical sequence of stages
  as a society progresses towards the ideal? Are they cyclical phases that
  societies go through over centuries or millennia? Or are they the three legs
  of a tripod that must always be kept in balance? 

  As we drink this third cup of wine, we reflect that whatever our personal
  views are, broadening our understanding of how other cultures view freedom
  will aid our interactions in our increasingly interconnected world.
}


\newcommand{\wineFourthCup}{
  \newReading As our seder draws near to the end, we take up our cups one last
  time as we think about the different views of freedom within America itself:

  \subsubsection*{The Two Types of Freedom}
  What I have found is that in the debates between left and right, people are
  arguing past each other rather than with each other. The left and the right
  both believe in freedom, but they just view it differently. In this sense
  there are two types of freedom. There is negative liberty or \textit{freedom
  from}, which is the main principle advocated by the right, and there is
  positive liberty or \textit{freedom to}, which is the main principle advocated
  by the left.

  Take a business for example. The right would argue that the owner should be
  free to run the business as they please. They shouldn't be restricted in who
  they can hire, how much they pay them, the conditions they work in etc. In
  this sense freedom means not having anyone else (namely the government)
  interfere with your business and impose restrictions on you. This is negative
  liberty\textemdash{}not because it is bad, but because it is freedom
  \textit{from} restrictions. In this view, the government is an oppressor that
  restricts the freedom of individuals. This is a very clear and obvious kind of
  freedom and is the most commonly discussed kind.

  \newReading But the left has its own view of freedom too. In this view people
  have a right to food, clothes, housing, education, healthcare, welfare,
  employment, equality etc. They should be free from poverty and discrimination.
  They should be free to live a full life. To a leftist, someone is not free if
  they cannot afford an education. They are not free if they cannot afford life
  saving healthcare. This is a far less clear and visible form of freedom and is
  far less often discussed. But just because the chains of poverty are hard to
  see, does not mean they are any less restrictive. In this view the government
  \textit{advances} freedom by allowing people to reach their full potential.

  There is one big difference between positive and negative liberty though;
  positive liberty costs money...a lot of money. Where do you draw the line so
  that the amount of spending is enough to guarantee positive liberty for all
  citizens? Can we ever reach a situation of full positive liberty or is that
  like chasing a rainbow?

  But clearly, negative liberty alone is not enough. It is not much good telling
  a starving man that he has freedom of speech or religion or is free to enter
  into a voluntary contract with a business without government interference. A
  homeless man is not restricted by any one person, but is still oppressed.

  \newReading The two types do not always line up perfectly with pro- or
  anti-government sentiments. Negative liberty requires the state to protect
  rights and enforce contracts. Even private property is dependent on the state.
   Likewise, the right would argue that the best way to achieve maximum positive
  liberty is through the free market. So neither type of freedom is
  intrinsically left or right, it depends on the circumstance.

  Freedom is all about doing what you want to do. But there are many complex
  layers to this deceptively simple idea. Freedom can mean avoiding government
  restrictions but it can also mean avoiding the invisible restrictions of
  poverty. To promote freedom and ensure everyone has a chance to reach their
  potential and fulfill their goals, we need to remove all the barriers, both
  obvious and hidden. Neither left nor right have a monopoly on freedom, just
  different views of it.

  \textemdash{}Adapted from postings by Robert Nielson

  With the fourth cup of wine, we think about which positive and negative
  liberties are most important in our own view of freedom.

}
